An efficient acyclic contact planner for multiped robots

### 2016-01-27 body/appendix_algo.tex
nicolas comments

Then, proving that $\mathbf{J}^{\dagger T} \mathbf{J}^{\dagger} = (\mathbf{J}\mathbf{J}^T)^{-1}$ is straight-forward,
using the SVD decomposition of each term.

>>>>

Then, the equality $\mathbf{J}^{\dagger T} \mathbf{J}^{\dagger} = (\mathbf{J}\mathbf{J}^T)^{-1}$ follows from the SVD decomposition of each term
~\citep{ben2003generalized}.

### 2016-01-27 body/efort.tex
nicolas comments

\subsubsection{The force and velocity ellipsoids.}
For a limb configuration $\mathbf{q}^k$, the Jacobian matrix
 $\mathbf{J}^k(\mathbf{q}^k)$   defines the relation:

@@

\subsubsection{Manipulability-based heuristics.}
From these definitions, we can derive three useful heuristics, that all account for the environment and the task being performed.
The first one, EFORT, was introduced by \cite{Tonneau2014}; the other two are new minor contributions, derived from these previous works.

@@

\begin{equation}
h_{EFORT}(\mathbf{q}, \mathbf{m}) = [\mathbf{m}^{T}(\mathbf{J}\mathbf{J}^T)\mathbf{m}]^{-\frac{1}{2}} ( \mu \mathbf{n}^T \mathbf{m})
\end{equation}
where $\mu$ and $\mathbf{n}$ are respectively the friction coefficient and the normal vector of the contact surface.

@@

$h_{EFORT}$ and $h_{vel}$ will favor contacts that allow large efforts or fast modifications in the velocity.
$EFORT$ in particular is useful for tasks such as standing up, pushing / pulling.
In other less demanding cases, manipulability can also be considered to avoid singularities.
To do so, we can consider the manipulability measure $h_{w}$, also given by \citeauthor{Yoshikawa1984}:

@@

$h_{EFORT}$, $h_{vel}$ and $h_{w}$ define three kinematic heuristics, fast to compute (indeed, thanks to our sampling-based approach, the Jacobian and inverse products
can be precomputed off-line), that allow the planner to select the best
candidates according to user-defined criterion.
%~ This results in faster computation times, and a reduced number of overall contact switches.

>>>>

\subsubsection{The force and velocity ellipsoids:}
For a limb configuration $\mathbf{q}^k$, the Jacobian matrix
 $\mathbf{J}^k(\mathbf{q}^k)$   defines the relation:

@@

\subsubsection{Manipulability-based heuristics:}
From these definitions, we can derive three useful heuristics, that all account for the environment and the task being performed.
The first one, EFORT, was introduced by \cite{Tonneau2014}; the other two are new minor contributions, derived from these previous works.

@@

\begin{equation}
h_{\textrm{\it EFORT}}(\mathbf{q}, \mathbf{m}) = [\mathbf{m}^{T}(\mathbf{J}\mathbf{J}^T)\mathbf{m}]^{-\frac{1}{2}} ( \mu \mathbf{n}^T \mathbf{m})
\end{equation}
where $\mu$ and $\mathbf{n}$ are respectively the friction coefficient and the normal vector of the contact surface.

@@

$h_{\textrm{\it EFORT}}$ and $h_{vel}$ will favor contacts that allow large efforts or fast modifications in the velocity.
$EFORT$ in particular is useful for tasks such as standing up, pushing / pulling.
In other less demanding cases, manipulability can also be considered to avoid singularities.
To do so, we can consider the manipulability measure $h_{w}$, also given by \citeauthor{Yoshikawa1984}:

@@

$h_{\textrm{\it EFORT}}$, $h_{vel}$ and $h_{w}$ define three kinematic heuristics, fast to compute (indeed, thanks to our sampling-based approach, the Jacobian and inverse products
can be precomputed off-line), that allow the planner to select the best
candidates according to user-defined criterion.
%~ This results in faster computation times, and a reduced number of overall contact switches.

### 2016-01-27 body/equilibrium.tex
nicolas comments

\subsubsection{Conditions for static equilibrium.}
We first define the variables of the problem, for $e$ contact points, expressed in world coordinates:
\begin{itemize}
\item $\mathbf{c} \in \mathbb{R}^3$ is the robot center of mass (COM);
%~ \item $\mathbf{L}  \in \mathbb{R}^3 $ is the angular momentum at the COM;
\item $\mathbf{m} \in \mathbb{R}$ is the robot mass;
\item $\mathbf{g} = [0,0,-9.81]^T$ is the gravity acceleration;
\item $\mu$ is the friction coefficient;
\item for the i-th contact point $1 \leq i \leq e$:
	\begin{itemize}
	\item $\mathbf{p_i}$ is the contact position;
	\item $\mathbf{f_i}$ is the force applied at $\mathbf{p_i}$;
	\item $\mathbf{n}_i,\mathbf{t}_{i1},\mathbf{t}_{i2}$ form a local Cartesian coordinate system centered at $\mathbf{p_i}$. $\mathbf{n}_i$ is aligned
	with the contact surface normal, and the $\mathbf{t}_i$s are tangent vectors.
	\end{itemize}
\end{itemize}

>>>>

\subsubsection{Conditions for static equilibrium:}
We first define the variables of the problem, for $e$ contact points, expressed in world coordinates:
\begin{itemize}
\item $\mathbf{c} \in \mathbb{R}^3$ is the robot center of mass (COM);
%~ \item $\mathbf{L}  \in \mathbb{R}^3 $ is the angular momentum at the COM;
\item $\mathbf{m} \in \mathbb{R}$ is the robot mass;
\item $\mathbf{g} = [0,0,-9.81]^T$ is the gravity acceleration;
\item $\mu$ is the friction coefficient;
\item for the i-th contact point $1 \leq i \leq e$:
	\begin{itemize}
	\item $\mathbf{p_i}$ is the contact position;
	\item $\mathbf{f_i}$ is the force applied at $\mathbf{p_i}$;
	\item $\mathbf{n}_i,\mathbf{t}_{i1},\mathbf{t}_{i2}$ form a local Cartesian coordinate system centered at $\mathbf{p_i}$. $\mathbf{n}_i$ is aligned
	with the contact surface normal, and the $\mathbf{t}_i$s are tangent vectors.
	\end{itemize}
\end{itemize}

### 2016-01-27 body/intro.tex
nicolas comments

As far as robotics applications are concerned, none of the existing planners is able to solve the problem with \gls{interactive} performances.
However, recent contributions to the interpolation between contact poses (problem $\mathcal{P}_3$) have brought promising preliminary solutions \citep{Hauser2014, herzog2015trajectory, Park116, Carpentier2016}. In particular, \citeauthor{Carpentier2016} are able to achieve this with \gls{interactive} performances on a real robot.
Therefore, a planner capable of solving \Pa and \Pb efficiently could outperform all existing planners if coupled with an interpolation method solving $\mathcal{P}_3$.
The main contribution of this paper is exactly this planner.

@@

\noindent \textbf{Regarding $\mathcal{P}_1$}, the guide path must guarantee the existence of a contact sequence to actuate it. We call this property \glslink{equilibrium feasible}{\textit{equilibrium feasibility}}. This property has not been studied yet; the only way to validate a waypoint in the path is to explicitly compute the contact locations and forces, which is computationally not reasonable \citep{Bouyarmane2009}, unless the scenario is limited to cyclic, quasi-flat cases \citep{zucker2010optimization}. \\

@@

\subsection*{Comparison with our previous work.}
The present paper is an extension of a conference paper to appear in the proceedings of the ISRR'15 conference~\citep{tonneauisrr15}.
The conference paper focuses on the theoretical formulation of the problem, and presents results obtained with virtual avatars.
This extension completes this work and brings new contributions, resulting from the implementation of our approach on real-world robots and problems.
In particular, a strong contribution is the introduction of a robust equilibrium criterion, designed to ensure the equilibrium of the robot despite bounded errors in the contact forces, coupled
with heuristics for relevant contact selection. These two contributions were required to apply our method to real-robot models, namely HRP-2 and HyQ~\citep{semini11hyqdesignjsce}.
In the present paper, a complete algorithm for the contact planner is given, and an access to the source code of our implementation is provided.
Furthermore, our experiments are supported by an in-depth analysis of our performances, as well as a discussion on the influence of the parameters to the method.

>>>>

As far as robotics applications are concerned, none of the existing planners is able to solve the problem with \gls{interactive} performances.
However, recent contributions to the interpolation between contact poses (problem $\mathcal{P}_3$) have brought promising preliminary solutions \citep{Hauser2014, herzog2015trajectory, Park116, Carpentier2016}. In particular, \citeauthor{Carpentier2016} are able to achieve this with \gls{interactive} performances on a real robot.
Therefore, a planner capable of efficiently solving \Pa and \Pb could outperform all existing planners if coupled with an interpolation method solving $\mathcal{P}_3$.
The main contribution of this paper is exactly this planner.

@@

\noindent \textbf{Regarding $\mathcal{P}_1$}, the guide path must guarantee the existence of a contact sequence to actuate it. We call this property \glslink{equilibrium feasible}{\textit{equilibrium feasibility}}. This property has not been studied yet; the only rigorous way to validate a waypoint in the path is to explicitly compute the contact locations and forces, which is computationally not reasonable \citep{Bouyarmane2009}, unless the scenario is limited to cyclic, quasi-flat cases \citep{zucker2010optimization}. \\

@@

\subsection*{Comparison with our previous work}
The present paper is an extension of a conference paper to appear in the proceedings of the ISRR'15 conference~\citep{tonneauisrr15}.
The conference paper focuses on the theoretical formulation of the problem, and presents results obtained with virtual avatars.
This extension completes this work and brings new contributions, resulting from the implementation of our approach on real-world robots and problems.
In particular, a strong contribution is the introduction of a robust equilibrium criterion, designed to ensure the equilibrium of the robot despite bounded errors in the contact forces, coupled
with heuristics for relevant contact selection. These two contributions were required to apply our method to real-robot models, namely HRP-2 and HyQ~\citep{semini11hyqdesignjsce}.
In the present paper, a complete algorithm for the contact planner is given, and an access to the source code of our implementation is provided.
Furthermore, our experiments are supported by an in-depth analysis of our performances, as well as a discussion on the influence of the parameters to the method.

### 2016-01-27 body/overview.tex
part 3 with andrea
nicolas comments

\subsection{Computation of a guide path --- \Pa}
We first consider the problem of planning a root guide path. 
%The objective is to compute a path of root configurations that will allow for equilibrium. 
%A desired requirement is to preserve the completeness of the planner: it should be able to explore any valid guide path, but at the same time, any computed guide path must lead to a sequence of configurations in static equilibrium.
\begin{figure}[t]
\centering
  \begin{overpic}[width=1\linewidth]{figures/2D_feas}
		\put (15,) {$C_{Equil}^0$      $\subset$} 
		\put (47,) {$C_{Contact}^0$ $\approx$ } 
		\put (76,) {$C_{Reach}^0$} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Illustration of several root configurations sets used in this paper in a 2D scene. Obstacles are violet, and units are in meters. To show the sets in a 2D representation, all the rotational joints of HRP-2 are locked in the shown configuration, such that a torso configuration
is only described by two positional parameters (x and y). The root of the robot is indicated with a black cross. To compute the reachable workspace, the point on the ankle indicated by a green cross was used. $C_{Equil}^0$ is included in $C_{Contact}^0$. $C_{Reach}^0$ approximates $C_{Contact}^0$. Depending on a parametrization, we can obtain $C_{Contact}^0 \subset C_{Reach}^0$. Considering the configurations around the top obstacle, we can observe a dramatic
divergence between  $C_{Equil}^0$  and $C_{Contact}^0$ when the problem is not \textit{\gls{cluttered}}.}
		   \label{fig:dedefeas}
\end{figure}
%
Ideally we would like to sample root configurations that are \gls{equilibrium feasible}. A root configuration is \gls{equilibrium feasible} if and only if there exists at least one joint configuration such that:
\begin{description}
\item[C1]: the robot is not in collision with the environment, 
%~ \item[C2] has at least one end-effector in contact with the environment, 
\item[C2]: the robot is in static equilibrium. 
\end{description}
Verifying C2 is computationally expensive because it requires the projection of the whole-body configuration into a contact posture, as well as the resolution of a linear program to compute the contact forces~\citep{Prete2016}.
For this reason, we are interested in finding an approximation of the \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} that is computationally efficient.
A requirement for C2 is the condition:
\begin{description}
\item[C3]: at least one end-effector is in contact with the environment.
\end{description}
C1 and C3 define \glslink{contact feasible}{\textit{contact feasibility}}, which is easier to verify than \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} because it does not require the computation of contact forces.
We make the assumption that for the \textit{\gls{cluttered} contact planning problem}, most of the times \gls{contact feasible} root configurations are also \gls{equilibrium feasible}.
For this reason we neglect C2 during the resolution of \Pa, and we consider it only when solving \Pb.
However, even verifying \glslink{contact feasible}{\textit{contact feasibility}} is too computationally expensive because it requires the projection of the whole-body configuration into a contact posture.

@@

@@

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing $R^0$ is denoted $W^0$.
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 The reachable workspace of a limb $R^k$ is denoted $W^k$: %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\}
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0 = \mathbf{0}$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define \mbox{$W = \bigcup_{k=1}^{l}W^k$}, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ for the root configuration $\mathbf{q}^{0}$.

>>>>

\subsection{Computation of a guide path --- \Pa}
We first consider the problem of planning a root guide path. 
%The objective is to compute a path of root configurations that will allow for equilibrium. 
%A desired requirement is to preserve the completeness of the planner: it should be able to explore any valid guide path, but at the same time, any computed guide path must lead to a sequence of configurations in static equilibrium.
\begin{figure}[t]
\centering
  \begin{overpic}[width=1\linewidth]{figures/2D_feas}
		\put (15,) {$C_{Equil}^0$      $\subset$} 
		\put (47,) {$C_{Contact}^0$ $\approx$ } 
		\put (76,) {$C_{Reach}^0$} 
		%~ \put (5,27) {3.b)} 
		%~ \put (37,27) {4.a)} 
		%~ \put (68,27) {4.b)} 
	\end{overpic}
\caption{Illustration of several root configurations sets used in this paper in a 2D scene. Obstacles are violet, and units are in meters. To show the sets in a 2D representation, all the rotational joints of HRP-2 are locked in the shown configuration, such that a torso configuration
is only described by two positional parameters (x and y). The root of the robot is indicated with a black cross. To compute the reachable workspace, the point on the ankle indicated by a green cross was used. $C_{Equil}^0$ is included in $C_{Contact}^0$. $C_{Reach}^0$ approximates $C_{Contact}^0$. Depending on a parametrization, we can obtain $C_{Contact}^0 \subset C_{Reach}^0$. Considering the configurations around the top obstacle, we can observe a dramatic
divergence between  $C_{Equil}^0$  and $C_{Contact}^0$ when the problem is not \textit{\gls{cluttered}}.}
		   \label{fig:dedefeas}
\end{figure}
%
Ideally we would like to sample root configurations that are \gls{equilibrium feasible}. A root configuration is \gls{equilibrium feasible} if and only if there exists at least one joint configuration such that:
\begin{description}
\item[C1]: the robot is not in collision with the environment, 
%~ \item[C2] has at least one end-effector in contact with the environment, 
\item[C2]: the robot is in static equilibrium. 
\end{description}
Verifying C2 is computationally expensive because it requires the projection of the whole-body configuration into a contact posture, as well as the resolution of a linear program to compute the contact forces~\citep{Prete2016}.
For this reason, we are interested in finding an approximation of the \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} that is computationally efficient.
A requirement for C2 is the condition:
\begin{description}
\item[C3]: at least one end-effector is in contact with the environment.
\end{description}
C1 and C3 define \glslink{contact feasible}{\textit{contact feasibility}}, which is easier to verify than \glslink{equilibrium feasible}{\textit{equilibrium feasibility}} because it does not require the computation of contact forces.
We make the assumption that for the \textit{\gls{cluttered} contact planning problem}, most of the times \gls{contact feasible} root configurations are also \gls{equilibrium feasible}.
For this reason we neglect C2 during the resolution of $\mathcal{P}_1$, and we consider it only when solving \Pb.
However, even verifying \glslink{contact feasible}{\textit{contact feasibility}} is computationally too expensive because it requires the projection of the whole-body configuration into a contact posture.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.95\linewidth]{figures/HyQ_roms}
  \caption{
           Reachable workspace and torso bounding box of HyQ. The green shapes represent the reachable workspace $W^k$ of each limb. The red shape is $W^0$.}
		   \label{fig:HyQ_roms}
\end{figure}

@@

%\subsubsection{Environment and other volumes} 
%\label{sec:A}
\medskip
The volume encompassing $R^0$ is denoted $W^0$ (Figure~\ref{fig:HyQ_roms}).
 %~ (Figure~\ref{fig:character}-right: central cylinder). 
 The reachable workspace of a limb $R^k$ is denoted $W^k$: %(Figure~\ref{fig:character}-right: the four ellipses). 
\begin{equation}
  W^k = \left\{ {\mathbf{x} \in \mathbb{R}^3: \exists \, \mathbf{q}^k \in C^k_{jl}, \mathbf{p}^k(\mathbf{q}^k) = \mathbf{x} } \right\}
\end{equation}
%~ \adnote{in this definition you don't account for joint limits}
where $\mathbf{p}^k$ denotes the end-effector position of $R^k$ (translation only) for $\mathbf{q}^0 = \mathbf{0}$ being the null displacement, and  $C^k_{jl}$ is the space
of admissible limb joint configurations. We also define \mbox{$W = \bigcup_{k=1}^{l}W^k$}, and
$W^k(\mathbf{q}^{0})$ (for $1 \leq k \leq l$) as the volume $W^k$ for the root configuration $\mathbf{q}^{0}$.

### 2016-01-27 body/rbprm.tex
nicolas comments

\subsubsection*{Contact feasibility: necessary condition}
For a contact to be possible, a volume $O_i \in O$ necessarily intersects with the reachable workspace $W(\mathbf{q}^{0})$ (Figure~\ref{fig:contact_gen}--1). Furthermore, if $\mathbf{q}^{0}$ is \gls{contact feasible}, then the torso of the robot $W^0(\mathbf{q}^{0})$ is necessarily not colliding  with the environment $O$.

Therefore we can define an outer approximation  $C^0_{Nec} \supset$ \gls{ccontact0} defined as: 
\begin{equation}
C_{Nec}^0 = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%~ The inclusion \gls{ccontact0} $\subset C^0_{Nec}$ is straightforward, but it is very important.
%~ Planning in $C^0_{Nec}$ allows a strong reduction of the search space, while not discarding 
%~ any root configuration that could lead to a solution.

%~ The condition defining $C^0_{Nec}$ is only necessary. This means that it might not be possible to extend a guide path planned 
%~ in $C^0_{Nec}$ into a sequence of contact configurations.

\subsubsection*{Contact feasibility: sufficient condition}
%~ In building $C_{reach}^1$, we directly considered the including hull of the root body $W^{0}$, obtaining a necessary condition. 
A trivial sufficient condition for \glslink{contact feasible}{\textit{contact feasibility}} can be constructed as a variation of $C^0_{Nec}$, by replacing $W^0$ with a bounding volume $B^{Suf}$ encompassing the whole robot in a given pose, except for the effector surfaces to be in contact. We denote by \mbox{$C_{Suf}^0 \subset $ \gls{ccontact0}} the set of root configurations corresponding to this sufficient condition:

\begin{equation}
C_{Suf}^0 = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } B^{Suf}(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}

%~ In general, the inclusion is strict, which means that we lose the completeness of the two-stage contact planner (i.e. the planner is not able to discover a path inside \mbox{$C_{reach} \setminus C_{Suf}^0$}). However, the sufficient condition guarantees that any such path leads to a valid sequence of contacts.

\subsection{Contact feasibility: a compromise \gls{reachc}}
The sufficient condition is not interesting in practice since it leads the solver to lose too many interesting paths. The necessary condition is not perfect either, since the first stage of the planner would stop on a guide that is not \gls{contact feasible}. 
An ideal shape $B$ (with $W^0 \subset B \subset B^{Suf}$) that leads to a necessary and sufficient condition may exist---even if it seems intuitively very unlikely in general. 
%~ Its construction is out of the scope of this work.

%~ It might be possible to find an ideal shape $B$ that leads to a necessary and sufficient condition; however, it seems intuitively very unlikely in general. The construction of a shape $W^0 \subset B \subset B^{Suf}$ leading to a necessary and sufficient condition (or the proof of its nonexistence) is out of the scope of this work.

However, using a shape between $W^0$ and $B^{Suf}$ leads to a trade-off between a necessary and a sufficient condition. We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$.
%
We then consider the spaces $C_{s}^0$
 \begin{equation}
C^0_s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
If $s=1$, then $W^0_s$ = $W^0$, such that $C_1^0$ = $C_{Nec}$. We thus consider that $s \geq 1$, since smaller values would only worsen the approximation.
By increasing $s$, the condition can become sufficient.
%~ The higher $s$, the closer this \gls{reachc} is to being sufficient, and if $s=1$, the search is complete (in the sense that all \gls{contact feasible} configurations can be found). 
The parametrization of $s$ then defines a trade-off between these two interesting extremes. 
We can choose $s$ by hand, or automatically as explained in Section~\ref{sec:params}.
The chosen value $s=s^*$ defines the \gls{reachc}, therefore we write \gls{creach0} $= C^0_{s^*}$.
%~ We call this condition the \gls{reachc}, which is defined by the chosen value $s=s^*$, and define \gls{creach0} $= C^0_{s^*}$.
%~ In the rest of the paper, we consider that an appropriate value $s^*$ of $s$ is chosen, and define \gls{creach0} $= C^0_{s^*}$.

>>>>

\subsubsection*{Contact feasibility, a necessary condition:}
For a contact to be possible, a volume $O_i \in O$ necessarily intersects with the reachable workspace $W(\mathbf{q}^{0})$ (Figure~\ref{fig:contact_gen}--1). Furthermore, if $\mathbf{q}^{0}$ is \gls{contact feasible}, then the torso of the robot $W^0(\mathbf{q}^{0})$ is necessarily not colliding  with the environment $O$.

Therefore we can define an outer approximation  $ \mathcal{C}^0_{\textrm{\it Nec}} \supset$ \gls{ccontact0} defined as: 
\begin{equation}
\mathcal{C}^0_{\textrm{\it Nec}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } W^0(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%~ The inclusion \gls{ccontact0} $\subset \mathcal{C}^0_{\textrm{\it Nec}}$ is straightforward, but it is very important.
%~ Planning in $\mathcal{C}^0_{\textrm{\it Nec}}$ allows a strong reduction of the search space, while not discarding 
%~ any root configuration that could lead to a solution.

%~ The condition defining $\mathcal{C}^0_{\textrm{\it Nec}}$ is only necessary. This means that it might not be possible to extend a guide path planned 
%~ in $\mathcal{C}^0_{\textrm{\it Nec}}$ into a sequence of contact configurations.

\subsubsection*{Contact feasibility, a sufficient condition:}
%~ In building $C_{reach}^1$, we directly considered the including hull of the root body $W^{0}$, obtaining a necessary condition. 
A trivial sufficient condition for \glslink{contact feasible}{\textit{contact feasibility}} can be constructed as a variation of $\mathcal{C}^0_{\textrm{\it Nec}}$, by replacing $W^0$ with a bounding volume $B^{\textrm{\it Suf}}$ encompassing the whole robot in a given pose, except for the effector surfaces to be in contact. We denote by \mbox{$\mathcal{C}^0_{\textrm{\it Suf}} \subset $ \gls{ccontact0}} the set of root configurations corresponding to this sufficient condition:

\begin{equation}
\mathcal{C}^0_{\textrm{\it Suf}} = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } B^{\textrm{\it Suf}}(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}

%~ In general, the inclusion is strict, which means that we lose the completeness of the two-stage contact planner (i.e. the planner is not able to discover a path inside \mbox{$C_{reach} \setminus \mathcal{C}^0_{\textrm{\it Suf}}$}). However, the sufficient condition guarantees that any such path leads to a valid sequence of contacts.

\subsection{Contact feasibility: a compromise \gls{reachc}}
The sufficient condition is not interesting in practice since it leads the solver to lose too many interesting paths. The necessary condition is not perfect either, since the first stage of the planner would stop on a guide that is not \gls{contact feasible}. 
An ideal shape $B$ (with $W^0 \subset B \subset B^{\textrm{\it Suf}}$) that leads to a necessary and sufficient condition may exist---even if it seems intuitively very unlikely in general. 
%~ Its construction is out of the scope of this work.

%~ It might be possible to find an ideal shape $B$ that leads to a necessary and sufficient condition; however, it seems intuitively very unlikely in general. The construction of a shape $W^0 \subset B \subset B^{\textrm{\it Suf}}$ leading to a necessary and sufficient condition (or the proof of its nonexistence) is out of the scope of this work.

However, using a shape between $W^0$ and $B^{\textrm{\it Suf}}$ leads to a trade-off between a necessary and a sufficient condition. We define $W^0_s$ as the volume $W^0$ subject to a scaling transformation by a factor $s \in \mathbb{R}^+$.
%
We then consider the spaces $C_{s}^0$
 \begin{equation}
C^0_s = \{ \mathbf{q}^0 : W(\mathbf{q}^{0}) \cap O \neq \emptyset \text{ and } W^0_s(\mathbf{q}^{0}) \cap O = \emptyset \} % \\ }
 %~ & \text{ and } & A_{torso}(\mathbf{q}^{root}) \cap W = \emptyset \}
\end{equation}
%
If $s=1$, then $W^0_s$ = $W^0$, such that $C_1^0$ = $\mathcal{C}^0_{\textrm{\it Nec}}$. We thus consider that $s \geq 1$, since smaller values would only worsen the approximation.
By increasing $s$, the condition can become sufficient.
%~ The higher $s$, the closer this \gls{reachc} is to being sufficient, and if $s=1$, the search is complete (in the sense that all \gls{contact feasible} configurations can be found). 
The parametrization of $s$ then defines a trade-off between these two interesting extremes. 
We can choose $s$ by hand, or automatically as explained in Section~\ref{sec:params}.
The chosen value $s=s^*$ defines the \gls{reachc}, therefore we write \gls{creach0} $= C^0_{s^*}$.
%~ We call this condition the \gls{reachc}, which is defined by the chosen value $s=s^*$, and define \gls{creach0} $= C^0_{s^*}$.
%~ In the rest of the paper, we consider that an appropriate value $s^*$ of $s$ is chosen, and define \gls{creach0} $= C^0_{s^*}$.
